Unsupervised Semantic Mapping for Healthcare Data Storage Schema

### 2021-07-25 ManuScript/IEEEAccess/Revision 1/source/ManuScript.tex
revision 1

\begin{abstract}
Data, information, and knowledge processing systems, in the domain of healthcare are currently plagued by the heterogeneity at various levels. Current solutions, have focused on developing a standard based, manual intervention mechanism, which requires a large amount of human resource and necessitate realignment of existing systems. State-of-the-art methodologies in the field of natural language processing and machine learning, can help to partially automate this process, reducing the resource requirements and providing a relatively good multi-class based classification algorithm. We present a novel methodology for bridging the gap between various healthcare data management solutions by leveraging the strength of transformer based machine learning models, to create mappings between the data elements. Additionally, the annotated data, collected against five medical schemas and labeled by 4 annotators is made available for helping future researchers. Our results indicate, that for biased, dependent multi-class text classification, RoBERTa shows the best performance, achieving a Cohen's kappa score of 0.47 and Matthews Correlation Coefficient (MCC) score of 0.48, with human annotated data.
\end{abstract}

@@

Healthcare Standards such as HL7 - Fast Healthcare Interoperability Resources (FHIR), and openEHR provide the foundations for storing and communicating medical data, through the use of well defined protocols. While Systematized Nomenclature of Medicine—Clinical Terms (Snomed-CT) \cite{snomedct_url} and Logical Observation Identifiers Names and Codes (LOINC) \cite{loinc} provide a standard definition for clinical terminologies and laboratory tests, respectively. Similarly Medical Logic Module (MLM) provides a standardized way for expressing medical knowledge. However, the plethora of standards necessitates the creation of bridging standards that can resolve the heterogeniety between the medical standards. Substantial effort has gone into this endeavor with the Clinical Information Modeling Initiative (openCIMI) \cite{CIMI2015} taking the lead in bridging the gap between HL7v3 and openEHR. Similarly, SNOMED CT and LOINC are working to resolve the redundancies between the two terminological standards since 2013. This healthcare interoperability solution follows a formal, albiet long process, which is greatly dependent on the human factor. However, the current healthcare scenario requires a quick solution to create a scaffolding of an interoperable bridge between various healthcare providers. It is also important to ensure that this scaffolding should be able to support the formal standardization processes of the future. In \cite{Satti2020}, we have presented the Ubiquitous Health Platform (UHP), which provides semantic reconciliation-on-read based data curation for resolving data interoperability between various schema. This methodology is based on the creation and management of schema maps, that can provide the framework for transforming a source schema into a target schema. 

@@

Martono et al. \cite{martono2017review} provided overview of previous studies related to linguistic approaches used for schema matching. Linguistic methods focused on finding strings and evaluate there similarity in different schemas. The string are normally normalized before to align both the strings before similarity comparison. The normalized strings are categories based on the information relatedness and element with similar category are compared using various similarity measure including  Jaro-distance, Lavenstein (edit-distance), and many more. Alwan et al. \cite{alwan2017survey} have summarized the techniques used in the literature for schemas and instances based schema matching. The information used for schema matching is categories into  schema information, instance and auxiliary information. Most of the searchers have used syntactic techniques (including n-gram, and regular expression), semantic techniques (including Latent Semantic Analysis, WordNet/Thesaurus and Google Similarity) for schema level and instance level matching to achieve the final goal of data/information interoperability.  Kersloot et al. \cite{kersloot2020natural} performed a comprehensive systematic review to evaluate Natural Language Processing (NLP) algorithms used for clinical text mapping onto ontological concepts. The findings of the studies were evaluated with respect to five categories; use of NLP algorithms, data used, validation and evaluation performed, result presentation, and generalization of results. The authors revealed that over one-fourth of the NLP algorithms used were not evaluated and have no validation. The systems that claimed generalization, were self evaluated and having no external validation.

@@

\section{Methodology}
\label{methodology}
Healthcare interoperability, with a focus on non-standard compliant medical schema is dependent on the generation and validation of schema maps as discussed above. To this end, the creation of a cohesive workflow is of utmost importance. In our earlier work \cite{Satti2020} we used maximum sequence identification and suffix tress based matching for syntactic matching of two distinct data schemas. This was followed by semantic concept enrichment and subsequently concept matching for creating rules in the form of schema maps. The simplified mapping functions, thereby created, provided a simple methodology for converting semi-structured medical data into an interpretable model form.
In our current methodology as visualized in Figure \ref{methodology}, we have utilized state-of-the-art natural language processing (NLP) techniques to extract the schema mapping rules from semi-structured data schemas. This methodology is based on identifying similarity between vector representations of two attributes, belonging to different medical schemas. 
Traditional NLP techniques such as Word2Vec are able to convert a word into an embedded vector, while Bidirectional Encoder Representations from Transformers (BERT) extracts an embedded vector from a sentence \cite{zhu2018exploring}. However, the terms forming the attribute names are bigger than a word (combination of multiple words) and smaller than a sentence. In order to resolve this problem, we extracted the set of suffixes from the terms forming the attribute names. The bidirectional nature of BERT, allows the creation of contextual embedded vectors, where each target word is affected by its neighboring words. Hence to convert the set of suffixes into a sentence, we collected the set of concepts corresponding to each suffix, from UMLS. This operation has two effects, firstly it is used to remove any suffix, which does not have a corresponding concept and secondly the extracted concepts are used to add context to each suffix and produce a contextual sentence. 
The following subsections provide the practical details for our methodology from schema acquisition to attribute name expansion, and finally schema map generation.

@@

A similar notion of data heterogeneity, in terms of medical data schema is evident across the healthcare domain. This is caused by various factors, including the lack of one all-encompassing, and universally applicable terminological standard and different normalization level for representing attributes.\\
In the former case, while SNOMED-CT provides a mechanism for identifying the standard codes for clinical terms and LOINC can be used for laboratory related terms, most attribute names are created based on the gut feeling of the database designer. Additionally, while these codes can be used to represent data instances, the data schema, achieves no benefit from the same.  Consider the terms ``name'' and ``patientName'', which refer to the same attribute of the patient entity.  However, since there is no standard way to represent this attribute, both are considered correct ($s_1$ and $s_3$ use the former representation, while $s_2$ and $s_5$ use the latter).\\
In the later case, differences in normalization cause semantic differences, due to which some data could be available in one schema but absent in others such as OpenEMR demographics identifying the patient's residential location using specific attributes like ``Address'', ``City'', ``State'', ``Postal Code'', ``Country'', and others. Similarly, ``EncounterDate'' from $s_5$ is semantically similar to ``BeginDate'' of ``openemr\_MedicalProblems'' table in $s_1$, ``AdmissionStartDate'' of ``LabsCorePopulatedTable'' in $s_2$, 
``time'' in ``Diagnosis'' table of $s_3$, and  ``dateOfAdmission'' in ``Diagnosis'' and ``CareHistory'' tables of $s_4$.
Finally, $s_1$ and $s_3$ have separate tables containing the medicinal prescription details, however the same details are unavailable in $s_2$, $s_4$, and $s_5$. Once again, this is not an incorrect behavior since this information, might not be a part of the context or the requirements for the EMR/EHR storage systems.
In fact, the change in context of the medical data storage system from the initial time of development to a later stage of collaborative processing systems, is the main cause of heterogeneity. In order to provide an interoperable solution, it is therefore necessary to enhance the semantics of each data attribute by its contextually equivalent sentence.

@@

Schema Maps provide an interoperable bridge between two medical systems ($s_i \wedge s_j$), by identifying the semantic relationship between their participating attributes. This identification is based on the similarity between the embedded vectors, of the semantically enriched sentences corresponding to each data attribute. While the embedded vectors can be generated using any methodology, we tested 11 methodologies with Word2Vec and 10 models based on BERT (including BERT base, BERT large, distilBert, and RoBERTa) further detailed in section \ref{experimentalSetup}, the large/STSb version of RoBERTa\cite{reimers-2019-sentence-bert}, provides the best results. The pair of embedded vectors thus produced are then used to calculate cosine similarity, which is based on the inverse cosine distance between them. For our classification, we used the raw results (unnormalized) of cosine similarity, which produces a score between -1, and 1. Cosine similarity score of 0 indicates orthogonal relationship between the two vectors, which in our scenario indicates that the two sentences, and by extension their attributes are not related to each other. -1 indicates inverse relationship between the attributes, while 1 indicates the two attributes are very much the same.
For producing our schema maps, we are interested in three types of relationships, ``equal'' (the two attributes are same), ``related'' (the two attributes are related to each other), and ``unrelated'' (no relationship between the attributes). In order to classify the similarity results into one of these three classes, we then calculated the best thresholds using Matthews Correlation coefficient (MCC) \cite{chicco2020advantages} for classifying each instance as ``equal'', ``related'', and ``unrelated''.

\begin{equation}
\resizebox{0.9\hsize}{!}{$%
	MCC = \frac{(TP.FN) - (FP.FN)}{\sqrt{(TP+FP)(TP+FN)(TN+FP)(TN+FN)}}\longrightarrow [-1,1]
	$%
}%
\end{equation}
MCC provides a fair measure of the ability with which a classifier can correctly predict both positive and negative instances \cite{chicco2020advantages}. MCC score of 0 represents random classification, however, with an increase in the number of true positives and true negatives MCC moves closer to 1. It also takes into account the false positives and false negatives, which shift the mcc score towards -1. This measurement is markedly different from accuracy that fails to account for imbalanced datasets and $F_1$ measure which is not affected by the true negative scores. As a result, MCC provides an acceptable alternate in our current scenario comprising of imbalanced dataset (largely in favour of class ``unrelated'') to measuring the true performance of the models, used for threshold selection and model evaluation. Finally on a test dataset we evaluated our multi-class classification approach using MCC and Cohen's Kappa coefficient ($\kappa$) \cite{gwet2014handbook} to identify the relationships between each pair of attributes.

\section{Experimental Setup}
\label{experimentalSetup}
In our earlier work \cite{Satti2020} $s_1$, $s_2$, and $s_5$ were used to generate over 115 million patient records, which are converted into a semi-structured form and stored in Hadoop Distributed File System (HDFS). We extended the same setup to create an additional 100,000 records, for 1000 patients with 3 medical fragments for $s_1$, $s_2$, and $s_4$, and 97 randomly selected and generated medical fragments amongst $s_1$, $s_2$, $s_3$, $s_4$ and $s_5$. These fragments follow various design elements supporting a variety of valid relational storage architectures. Such as, $s_1$, $s_2$ and $s_4$ are represented by creating a separate medical fragment for each participating table, $s_3$ utilizes its medical fragment to generate a linked record (from a linked object graph), where by the attributes can refer to other objects besides the elements of $t$, mimicking the application of explicit foreign keys, and $s_5$ is a flat table structure. The code to generate this data set is available at ``uhp\_map\_generation''\footnote{https://github.com/desertzebra/UHP\_v4/tree/main/uhpr\_storage}. This application produces three custom formatted files containing an index for patients, an index for their medical fragments, and the medical fragment corresponding to the EMR data. Using the medical fragments file, we then generate the semantically enriched attribute \footnote{https://github.com/desertzebra/UHP\_v4/tree/main/uhp\_map\_generation}, which contains the suffixes and their concepts corresponding to each EMR data attribute. The resulting set of enriched attributes are temporarily stored in a JSON file, which is then read by the same application to partially generate the schema maps. This process is used to create 20,349 distinct pairs of attributes across $s$. Each pair also contains the ``relationshipList'', which stores the results of fuzzy string matching\cite{FuzzyWuzzy} \footnote{Java Library: https://github.com/xdrop/fuzzywuzzy} between the attribute names. 
The JSON file thus produced is then used by a python script to generate the semantically enriched sentences and their embedded vectors using Word2Vec, and 10 pre-trained BERT NLI models \cite{reimers-2019-sentence-bert}. The BERT models include ``bert-base-nli-stsb-mean-tokens'', ``bert-large-nli-stsb-mean-tokens'', ``roberta-base-nli-stsb-mean-tokens'', ``roberta-large-nli-stsb-mean-tokens'', ``distilbert-base-nli-stsb-mean-tokens'', ``bert-base-nli-mean-tokens'', ``bert-large-nli-mean-tokens'', ``roberta-base-nli-mean-tokens'', ``roberta-large-nli-mean-tokens'', and ``distilbert-base-nli-mean-tokens''. 
The embedding vectors are then compared using cosine similarity, which produces a score between -1 and 1.
The rationale behind switching the applications at various stages is to cache the results and create checkpoints for restarting any failed stages, easily. Additionally, since python provides better support for easy generation of embedding vectors, it was thus preferred over the Java based implementation, which is otherwise very beneficial for other tools.
%These applications were executed on a workstation running Ubuntu 20.04.2 on top of AMD Ryzen 3 2200G, and 32GB ram.%

@@

\subsection{Dataset Annotation}
In order to compare our computed models with ground truth and to identify the best thresholds for classifying each instance as ``equal'', ``related'', or ``unrelated'' 4 human annotators were utilized to anonymously, score the similarity of each pair of attribute names. In order to support this process, we first repurposed one of our generated data matrix by marking all attribute pairs belonging to the same schema with the symbol ``-''. Following this, the annotators marked each cell corresponding to a pair of attributes (conversely, each attribute pair corresponds to two cells with the positioning of the pair-participants swapped; which is used for clarity and identify correct relationships between the attribute on left and attribute on right), by determining the similarity in terms of dissimilar as ``0'', exactly similar as ``1'', row attribute as child of column attribute as ``<'', row attribute as a parent of the column attribute as ``>'', and finally, unknown as ``~''. The data sheets generated after this extensive human effort have been made available for other researchers\footnote{\url{https://github.com/desertzebra/EMR-Interoperability/tree/master/Implemenation/Data/Annotated}}.

@@

These sheets, additionally contain some missing values, which were left out by the annotators but in order to maintain their originality, these values were not filled; instead during our evaluation for these datasets, the missing values were considered as having the score ``0''. Using $\kappa$, we evaluated the inter-rater agreement of these annotations, which have been visualized in Figure \ref{fig:kappaInterAnnotatorAgrement}. It can be seen in this plot, that ``Annotator3'' has very small correlation with the other 3 annotators. This difference can be traced back to the number and type of annotations performed by each annotator, which is shown in Table \ref{tab:annotator_marking}. The ``Annotator3'' has marked 2103 cells as related (one of >, < , or ~) and left 153 as empty. However, even in the presence of these difference, it is pertinent to include the data for all annotators in order to avoid any biasedness.

@@

\subsection{Threshold Selection}
A good text classification methodology is dependent on the correct choice of a threshold, which can maximize the target class participation. In case of independent labels, area under the precision recall curve can provide this optimal measure, however as in our case, for dependent classes on a biased dataset the MCC, is better\cite{chicco2020advantages}.
Since our aim is to apply an optimal text similarity classifier to resolve this multi-class problem ($class_0$, $class_{0.5}$, and $class_1$), we have to test various threshold scores for separating the instances between $class_0$ and $class_{0.5}$ ($t_1$), and then $class_{0.5}$ and $class_1$ ($t_2$). Additionally, since our aim is to correctly identify the similar attribute instances, it is pertinent to maximize the classification performance of $class_1$ (similar), followed by $class_{0.5}$ (related), and finally $class_0$ (unrelated).
With a step size of \textit{0.05} ($step$), and starting from $t_1$ as 0.0 and $t_2$ as $t_1$ + $step$, we move the thresholds until $t_2$ reaches \textit{1.0}, followed by increase in $t_1$ by step size. Eventually, $t_1$, reaches \textit{0.95} and $t_2$ reaches \textit{1.0}, at which point, the process stops. This is to ensure that $t_1$ remains behind $t_2$, for all iterations, measuring MCC score, for the 12 models. These models include, ``Fuzzy\_Wuzzy'', ``Word2Vec'', and 10 BERT based models.

@@

\subsection{Model Evaluation}
On unseen test dataset with thresholds selected in the previous step and the 12 models, we measured the performance score using one vs all binarization of the multi-classes. As evident in Table \ref{tab:evaluations}, very high values of accuracy are visible across all models with all three positive classes. In all, except the case of Word2Vec, Precision and recall also show values close to \textit{1.0}. However, these measures are very misleading, since the dataset is greatly biased in favour of $class_0$. In terms of $F_1$ measure $class_{0.5}$ shows the worst possible results, independently, with all except Word2Vec and bert-large-nli-mean-tokens having a score of \textit{0.0}. Bert-large-nli-mean-tokens provides the best $F_1$ measure at \textit{0.49}. These metrics are thus not useful to gauge the performance of the evaluated models. Instead, focusing on the MCC score, provides a good picture of the model performance for individual classes when all other instances are negative.

Finally, we evaluated the overall $\kappa$ coefficient and MCC score to evaluate the performance of each model on the test dataset. These scores range between [-1,1], providing a measure quantifying the accuracy of the classifier to correctly predict correct and incorrect instances. The models Word2Vec, bert-large-nli-mean-tokens, and roberta-base-nli-mean-tokens, with values between [0,0.20] indicate random classification, while $\kappa$ score between [0.21,0.39], achieved by fuzzy wuzzy, bert-base-nli-mean-tokens, roberta-base-nli-sts-mean-tokens, distilbert-base-nli-stsb-mean-tokens, bert-base-nli-mean-tokens, roberta-large-nli-mean-tokens, and distilbert-base-nli-mean-tokens, show only minimal agreement with the annotated data \cite{mchugh2012interrater}.  Relative best rates, in this case, are achieved by roberta-large-nli-stsb-mean-tokens, which substantially surpasses the other models, however in absolute terms, it shows only weak agreement. Similarly, MCC scores show a similar result, with roberta-large-nli-stsb-mean-tokens, achieving the relative best performance amongst all classifiers.

>>>>

\begin{abstract}
Data, information, and knowledge processing systems, in the domain of healthcare are currently plagued by the heterogeneity at various levels. Current solutions, have focused on developing a standard based, manual intervention mechanism, which requires a large amount of human resource and necessitate realignment of existing systems. State-of-the-art methodologies in the field of natural language processing and machine learning, can help to partially automate this process, reducing the resource requirements and providing a relatively good multi-class based classification algorithm. We present a novel methodology for bridging the gap between various healthcare data management solutions by leveraging the strength of transformer based machine learning models, to create mappings between the data elements. Additionally, the annotated data, collected against five medical schemas and labeled by four annotators is made available for helping future researchers. Our results indicate, that for biased, dependent multi-class text classification, transformer based models provide better results than linguistic and other classical models. In particular, the Robustly Optimized BERT Pretraining Approach (RoBERTa) provides the best schema matching performance by achieving a Cohen's kappa score of 0.47 and Matthews Correlation Coefficient (MCC) score of 0.48, with human annotated data.
\end{abstract}

@@

\\
Healthcare Standards such as HL7 - Fast Healthcare Interoperability Resources (FHIR), and openEHR provide the foundations for storing and communicating medical data, through the use of well defined protocols. While Systematized Nomenclature of Medicine—Clinical Terms (Snomed-CT) \cite{snomedct_url} and Logical Observation Identifiers Names and Codes (LOINC) \cite{loinc} provide a standard definition for clinical terminologies and laboratory tests, respectively. Similarly Medical Logic Module (MLM) provides a standardized way for expressing medical knowledge. Variety in these and many other healthcare standards necessitates the creation of bridging standards that can resolve the heterogeneity between the medical standards \cite{mohammed2010ubiquitous}. Substantial effort has gone into this endeavor with the Clinical Information Modeling Initiative (openCIMI) \cite{CIMI2015} taking the lead in bridging the gap between HL7v3 and openEHR. Similarly, SNOMED CT and LOINC are working to resolve the redundancies between the two terminological standards since 2013. This healthcare interoperability solution follows a formal, albeit long process, which is greatly dependent on the human factor. However, the current healthcare scenario requires a quick solution to create a scaffolding of an interoperable bridge between various healthcare providers. It is also important to ensure that this scaffolding should be able to support the formal standardization processes of the future. In \cite{Satti2020}, we have presented the Ubiquitous Health Platform (UHP), which provides semantic reconciliation-on-read based data curation for resolving data interoperability between various schema. This methodology is based on the creation and management of schema maps, that can provide the framework for transforming a source schema into a target schema. 
\\

@@

Martono et al. \cite{martono2017review} provided an overview of a linguistic approach for schema matching. The authors presented various linguistic methods to identify token strings in element names, followed by similarity evaluation between various schema. The process starts by normalizing the strings, which can be achieved via tokenization, generalization, elimination or semantic tagging. The normalized strings are then categorized based on their information relatedness. Elements belonging to the same categories are then compared with each other using two similarity measures, which include Lavenstein distance between words and Jaro-distance between 3-grams character sub-strings. Alwan et al. \cite{alwan2017survey} has summarized the techniques used in literature for schema matching based on database schemas and instances. These techniques can be categorized based on the type of information used for schema matching which includes  schema level, instance level, hybrid (schema and instances) and auxiliary (which can include information from external sources). Accordingly, most research is focused towards schema level and instance level approaches which can utilize syntactic techniques (such as n-gram, and/or regular expressions) and/or semantic techniques (such as Latent Semantic Analysis, WordNet/Thesaurus, and Google Similarity), to achieve data/information interoperability.  Kersloot et al. \cite{kersloot2020natural} performed a comprehensive systematic review to evaluate Natural Language Processing (NLP) algorithms used for clinical text mapping onto ontological concepts. The authors categorized the findings of various studies based on the use of NLP algorithms, data, validation and evaluation techniques, result presentation, and generalization of results. The authors revealed that over one-fourth of the NLP algorithms used were not evaluated and have no validation. The systems that claimed generalization, were self evaluated and having no external validation.

@@

\section{Methodology}
\label{methodology}
Healthcare interoperability, with a focus on non-standard compliant medical schema is dependent on the generation and validation of schema maps as discussed above. To this end, the creation of a cohesive workflow is of utmost importance. In our earlier work \cite{Satti2020} we used maximum sequence identification and suffix trees based matching for syntactic matching of two distinct data schemas. This was followed by semantic concept enrichment and subsequently concept matching for creating rules in the form of schema maps. The simplified mapping functions, thereby created, provided a simple methodology for converting semi-structured medical data into an interpretable model form.
In our current methodology as visualized in Figure \ref{fig:methodology}, we have utilized state-of-the-art natural language processing (NLP) techniques to extract the schema mapping rules from semi-structured data schemas. This methodology is based on identifying similarity between vector representations of two attributes, belonging to different medical schemas. 
Traditional NLP techniques such as Word2Vec are able to convert a word into an embedded vector, while Bidirectional Encoder Representations from Transformers (BERT) extracts an embedded vector from a sentence \cite{zhu2018exploring}. Many of the attribute names within data schemas are represented by terms that are bigger than a word (combination of multiple words) and smaller than a sentence. In order to resolve this problem, we extracted the set of suffixes from the terms forming the attribute names. The bidirectional nature of BERT, allows the creation of contextual embedded vectors, where each target word is affected by its neighboring words. Hence to convert the set of suffixes into a sentence, we collected the set of concepts corresponding to each suffix, from UMLS. This operation has two effects, firstly it is used to remove any suffix, which does not have a corresponding concept and secondly the extracted concepts are used to add context to each suffix and produce a contextual sentence. 
The following subsections provide the practical details for our methodology from schema acquisition to attribute name expansion, and finally schema map generation.

@@

A similar notion of data heterogeneity, in terms of medical data schema is evident across the healthcare domain. This is caused by various factors, including the lack of one all-encompassing, and universally applicable terminological standard and different normalization level for representing attributes.\\
In the former case, while SNOMED-CT provides a mechanism for identifying the standard codes for clinical terms and LOINC can be used for laboratory related terms, most attribute names are created based on the gut feeling of the database designer. Additionally, while these codes can be used to represent elements in the data instances (such as when recording the disease name, a standard code is more beneficial than the text string for semantic interoperability), the elements in a data schema (such as attribute names which are used in queries) achieve no benefit from the same.  Consider the terms ``name'' and ``patientName'', which refer to the same attribute of the patient entity.  However, since there is no standard way to represent this attribute, both are considered correct ($s_1$ and $s_3$ use the former representation, while $s_2$ and $s_5$ use the latter).\\
In the later case, differences in normalization cause semantic differences, due to which some data could be available in one schema but absent in others such as OpenEMR demographics identifying the patient's residential location using specific attributes like ``Address'', ``City'', ``State'', ``Postal Code'', ``Country'', and others. Similarly, ``EncounterDate'' from $s_5$ is semantically similar to ``BeginDate'' of ``openemr\_MedicalProblems'' table in $s_1$, ``AdmissionStartDate'' of ``LabsCorePopulatedTable'' in $s_2$, 
``time'' in ``Diagnosis'' table of $s_3$, and  ``dateOfAdmission'' in ``Diagnosis'' and ``CareHistory'' tables of $s_4$.
Finally, $s_1$ and $s_3$ have separate tables containing the medicinal prescription details, however the same details are unavailable in $s_2$, $s_4$, and $s_5$. Once again, this is not an incorrect behavior since this information, might not be a part of the context or the requirements for the EMR/EHR storage systems.
In fact, the change in context of the medical data storage system from the initial time of development to a later stage of collaborative processing systems, is the main cause of heterogeneity. In order to provide an interoperable solution, it is therefore necessary to enhance the semantics of each data attribute by its contextually equivalent sentence.

@@

Schema Maps provide an interoperable bridge between two medical systems ($s_i \wedge s_j$), by identifying the semantic relationship between their participating attributes. This identification is based on the similarity between the embedded vectors, of the semantically enriched sentences corresponding to each data attribute. While the embedded vectors can be generated using any methodology, we tested 11 methodologies with Word2Vec and 10 models based on BERT further detailed in section \ref{experimentalSetup}. Our results indicate that the large/STSb version of Robustly Optimized BERT Pretraining Approach (RoBERTa)\cite{reimers-2019-sentence-bert}, provides the best matching results. The pair of embedded vectors thus produced are then used to calculate cosine similarity, which is based on the inverse cosine distance between them. For our classification, we used the raw results (unnormalized) of cosine similarity, which produces a score between -1, and 1. Cosine similarity score of 0 indicates orthogonal relationship between the two vectors, which in our scenario indicates that the two sentences, and by extension their attributes are not related to each other. -1 indicates inverse relationship between the attributes, while 1 indicates the two attributes are very much the same.
For producing our schema maps, we are interested in three types of relationships, ``equal'' (the two attributes are same), ``related'' (the two attributes are related to each other), and ``unrelated'' (no relationship between the attributes). In order to classify the similarity results into one of these three classes, we then calculated the best thresholds using Matthews Correlation coefficient (MCC) \cite{chicco2020advantages} for classifying each instance as ``equal'', ``related'', and ``unrelated''.

\begin{equation}
\label{eq:mccformula}
\resizebox{0.9\hsize}{!}{$%
	MCC = \frac{(TP \times TN) - (FP \times FN)}{\sqrt{(TP+FP)\times(TP+FN)\times(TN+FP)\times(TN+FN)}}\longrightarrow [-1,1]
	$%
}%
\end{equation}
MCC provides a fair measure of the ability with which a classifier can correctly predict both positive and negative instances \cite{chicco2020advantages}. The formula for calculating MCC is shown in Equation \ref{eq:mccformula}, which is based on classification performance measures such as true positive (TP), true negative (TN), false positive (FP), and false negative (FN). MCC score of 0 represents random classification, however, with an increase in the number of true positives and true negatives MCC moves closer to 1. It also takes into account the false positives and false negatives, which shift the MCC score towards -1. This measurement is markedly different from accuracy that fails to account for imbalanced datasets and $F_1$ measure which is not affected by the true negative scores. As a result, MCC provides an acceptable alternate in our current scenario comprising of imbalanced dataset (largely in favour of class ``unrelated'') to measuring the true performance of the models, used for threshold selection and model evaluation. Finally on a test dataset we evaluated our multi-class classification approach using MCC and Cohen's Kappa coefficient ($\kappa$) \cite{gwet2014handbook} to identify the relationships between each pair of attributes.

\section{Experimental Setup}
\label{experimentalSetup}
In our earlier work \cite{Satti2020} $s_1$, $s_2$, and $s_5$ were used to generate over 115 million patient records, which are converted into a semi-structured form and stored in Hadoop Distributed File System (HDFS). We extended the same setup to create an additional 100,000 records, for 1000 patients with 3 medical fragments for $s_1$, $s_2$, and $s_4$, and 97 randomly selected and generated medical fragments amongst $s_1$, $s_2$, $s_3$, $s_4$ and $s_5$. These fragments follow various design elements supporting a variety of valid relational storage architectures. Such as, $s_1$, $s_2$ and $s_4$ are represented by creating a separate medical fragment for each participating table, $s_3$ utilizes its medical fragment to generate a linked record (from a linked object graph), where by the attributes can refer to other objects, mimicking the application of explicit foreign keys, and $s_5$ is a flat table structure. The code to generate this data set is available at ``uhp\_map\_generation''\footnote{https://github.com/desertzebra/UHP\_v4/tree/main/uhpr\_storage}. This application produces three custom formatted files containing an index for patients, an index for their medical fragments, and the medical fragment corresponding to the EMR data. Using the medical fragments file, we then generate the semantically enriched attribute \footnote{https://github.com/desertzebra/UHP\_v4/tree/main/uhp\_map\_generation}, which contains the suffixes and their concepts corresponding to each EMR data attribute. The resulting set of enriched attributes are temporarily stored in a JSON file, which is then read by the same application to partially generate the schema maps. This process is used to create 20,349 distinct pairs of attributes across $s$. Each pair also contains the ``relationshipList'', which stores the results of fuzzy string matching\cite{FuzzyWuzzy} \footnote{Java Library: https://github.com/xdrop/fuzzywuzzy} between the attribute names. 
The JSON file thus produced is then used by a python script to generate the semantically enriched sentences and their embedded vectors using Word2Vec, and 10 pre-trained BERT NLI models \cite{reimers-2019-sentence-bert}. The BERT models include ``bert-base-nli-stsb-mean-tokens'', ``bert-large-nli-stsb-mean-tokens'', ``roberta-base-nli-stsb-mean-tokens'', ``roberta-large-nli-stsb-mean-tokens'', ``distilbert-base-nli-stsb-mean-tokens'', ``bert-base-nli-mean-tokens'', ``bert-large-nli-mean-tokens'', ``roberta-base-nli-mean-tokens'', ``roberta-large-nli-mean-tokens'', and ``distilbert-base-nli-mean-tokens''. 
The embedding vectors are then compared using cosine similarity. The rationale behind switching the applications at various stages is to cache the results and create checkpoints for restarting any failed stages, easily. Additionally, since python provides better support for easy generation of embedding vectors, it was thus preferred over the Java based implementation, which is otherwise very beneficial for other tools.
%These applications were executed on a workstation running Ubuntu 20.04.2 on top of AMD Ryzen 3 2200G, and 32GB ram.%

@@

\subsection{Dataset Annotation}
In order to compare our computed models with ground truth and to identify the best thresholds for classifying each instance as ``equal'', ``related'', or ``unrelated'' four human annotators were utilized to anonymously, score the similarity of each pair of attribute names. In order to support this process, we first repurposed one of our generated data matrix by marking all attribute pairs belonging to the same schema with the symbol ``-''. Following this, the annotators marked each cell corresponding to a pair of attributes (conversely, each attribute pair corresponds to two cells with the positioning of the pair-participants swapped; which is used for clarity and identify correct relationships between the attribute on left and attribute on right), by determining the similarity in terms of dissimilar as ``0'', exactly similar as ``1'', row attribute as child of column attribute as ``<'', row attribute as a parent of the column attribute as ``>'', and finally, unknown as ``~''. The data sheets generated after this extensive human effort have been made available for other researchers\footnote{\url{https://github.com/desertzebra/EMR-Interoperability/tree/master/Implemenation/Data/Annotated}}.

@@

These sheets, additionally contain some missing values, which were left out by the annotators but in order to maintain their originality, these values were not filled; instead during our evaluation for these datasets, the missing values were considered as having the score ``0''. Using $\kappa$, we evaluated the inter-rater agreement of these annotations, which have been visualized in Figure \ref{fig:kappaInterAnnotatorAgrement}. It can be seen in this plot, that ``Annotator3'' has very small correlation with the other 3 annotators. This difference can be traced back to the number and type of annotations performed by each annotator, which is shown in Table \ref{tab:annotator_marking}. The ``Annotator3'' has marked 2103 cells as related (one of >, < , or ~) and left 153 as empty. Even in the presence of these differences, it is pertinent to include the data for all annotators in order to avoid any bias.

@@

\subsection{Threshold Selection}
A good text classification methodology is dependent on the correct choice of a threshold, which can maximize the target class participation. In case of independent labels, area under the precision recall curve can provide this optimal measure, however as in our case, for dependent classes on a biased dataset the MCC, is better\cite{chicco2020advantages}.
Since our aim is to apply an optimal text similarity classifier to resolve this multi-class problem ($class_0$, $class_{0.5}$, and $class_1$), we have to test various threshold scores for separating the instances between $class_0$ and $class_{0.5}$ ($t_1$), and then $class_{0.5}$ and $class_1$ ($t_2$). Additionally, since our aim is to correctly identify the similar attribute instances, it is pertinent to maximize the classification performance of $class_1$ (similar), followed by $class_{0.5}$ (related), and finally $class_0$ (unrelated).
With a step size of \textit{0.05} ($step$), and starting from $t_1$ as 0.0 and $t_2$ as $t_1$ + $step$, we move the thresholds until $t_2$ reaches \textit{1.0}, followed by increase in $t_1$ by step size. Eventually, $t_1$, reaches \textit{0.95} and $t_2$ reaches \textit{1.0}, at which point, the process stops. This is to ensure that $t_1$ remains behind $t_2$, for all iterations, measuring MCC score, for the 12 models. These models include, ``Fuzzy\_Wuzzy'', ``Word2Vec'', and 10 BERT based models. The optimal thresholds achieved by each of these models is shown in Figure \ref{fig:thresholdSelection}.

@@

\subsection{Model Evaluation}
On unseen test dataset with thresholds selected in the previous step and the 12 models, we measured the performance score using one vs all binarization of the multi-classes. As evident in Table \ref{tab:evaluations}, very high values of accuracy are visible across all models with all three positive classes. In all, except the case of Word2Vec, precision and recall also show values close to \textit{1.0}. However, these measures are very misleading, since the dataset is greatly biased in favour of $class_0$. In terms of $F_1$ measure $class_{0.5}$ shows the worst possible results, independently, with all except Word2Vec and bert-large-nli-mean-tokens having a score of \textit{0.0}. Bert-large-nli-mean-tokens provides the best $F_1$ measure at \textit{0.49}. These metrics are thus not useful to gauge the performance of the evaluated models. Instead, focusing on the MCC score, provides a good picture of the model performance for individual classes when all other instances are negative.

Finally, we evaluated the overall $\kappa$ coefficient and MCC score to evaluate the performance of each model on the test dataset. These scores range between [-1,1], providing a measure quantifying the accuracy of the classifier to correctly predict correct and incorrect instances. As shown in Figure \ref{fig:evaluation}, the models Word2Vec, bert-large-nli-mean-tokens, and roberta-base-nli-mean-tokens, with values between [0,0.20] indicate random classification, while $\kappa$ score between [0.21,0.39], achieved by fuzzy wuzzy, bert-base-nli-mean-tokens, roberta-base-nli-sts-mean-tokens, distilbert-base-nli-stsb-mean-tokens, bert-base-nli-mean-tokens, roberta-large-nli-mean-tokens, and distilbert-base-nli-mean-tokens, show only minimal agreement with the annotated data \cite{mchugh2012interrater}.  Relative best rates, in this case, are achieved by roberta-large-nli-stsb-mean-tokens, which substantially surpasses the other models, however in absolute terms, it shows only weak agreement. Similarly, MCC scores show a similar result, with roberta-large-nli-stsb-mean-tokens, achieving the relative best performance amongst all classifiers.
